---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# urbnmapr

[![Travis-CI Build Status](https://travis-ci.org/UrbanInstitute/urbnmapr.svg?branch=master)](https://travis-ci.org/UrbanInstitute/urbnmapr)

The `urbnmapr` package provides state and county shapefiles in `tibble` format that is compatible to map with `ggplot2`.

Shapefiles include Alaska and Hawaii, transformed to be displayed as insets within the continental United States.

This package is heavily inspired by and derived in part from the [fiftystater package](https://cran.r-project.org/package=fiftystater) by William Murphy. In contrast, `urbnmapr`:

  - Uses shapefiles from the US Census Bureau
  - Converts the shapefile data to a `tibble` dataframe
  - Adds various identifiers for merging
  - Includes a county-level shapefile 

## Installation

You can install the latest version of `urbnmapr` from GitHub:

```{r installation, message=FALSE, warning=FALSE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("UrbanInstitute/urbnmapr")
```

## Usage

`urbnmapr` contains two `tibble` dataframes:

  - `states`
  - `counties`

The `states` and `counties` tibbles can be used with `geom_polygon()` and `coord_map()` to create base maps of the continental United States, with Alaska and Hawaii displayed as insets:

```{r blank-state, message=FALSE}
library(tidyverse)
library(urbnmapr)

states %>%
  ggplot(aes(long, lat, group = group)) +
    geom_polygon(fill = "grey", color = "#ffffff", size = 0.25) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

```{r blank-county}
counties %>%
  ggplot(aes(long, lat, group = group)) +
    geom_polygon(fill = "grey", color = "#ffffff", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

## Merging Data

The `states` and `counties` tibbles include various identifiers to simplify merging data. The states `states` tibble contains `state_fips`, `state_abbv`, and `state_name`. The `counties` tibble contains `county_fips`, `state_abbv`, `state_fips`,  `county_name`, and `state_name`. Both tibbles can be piped into `ggplot2` to create a choropleth map.

```{r us-choropleth}
statedata %>% 
  left_join(states, by = "state_name") %>% 
  ggplot(mapping = aes(long, lat, group = group, fill = horate)) +
  geom_polygon(color = "#ffffff", size = .25) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Homeownership rate")
  
```


```{r county}
household_data <- left_join(countydata, counties, by = "county_fips") 

household_data %>%
  ggplot(aes(long, lat, group = group, fill = medhhincome)) +
    geom_polygon(color = "#ffffff", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

## Styles

`library(urbnmapr)` works well with the [Urban Institute `ggplot2` theme.](https://github.com/UrbanInstitute/urban_R_theme)

```{r load-theme, message=FALSE}
source("https://raw.githubusercontent.com/UrbanInstitute/urban_R_theme/urban_R_theme_revamp/urban_theme.R")
```

```{r theme-state}

statedata %>% 
  left_join(states, by = "state_name") %>% 
  ggplot(mapping = aes(long, lat, group = group, fill = horate)) +
  geom_polygon(color = "#ffffff", size = .25) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradientn(labels = scales::percent) +
  urban_map +
   theme(legend.position = "right",
          legend.direction = "vertical",
          legend.title = element_text(face = "bold", size = 11),
          legend.key.height = unit(.2, "in")) +
  labs(fill = "Homeownership rate")

```

```{r theme-counties}
household_data %>%
  filter(state_name %in% c("Virginia", "Maryland", "District of Columbia")) %>%
  ggplot(aes(long, lat, group = group, fill = medhhincome)) +
    geom_polygon(color = "#ffffff", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    scale_fill_gradientn(labels = scales::dollar) +
    urban_map +
    theme(legend.position = "right",
          legend.direction = "vertical",
          legend.title = element_text(face = "bold", size = 11),
          legend.key.height = unit(.25, "in")) +
  labs(fill = "Median household income")
```

## License

Code released under the GNU General Public License v3.0.
