---
title: "Introducing urbnmapr"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
options(scipen = 999)
```

The `urbnmapr` package provides state, county, and other shapefiles in `tibble` format 
that is compatible to map with `ggplot2`. Shapefiles include Alaska and Hawaii, transformed to be displayed as insets 
within the continental United States.

This package is heavily inspired by and derived in part from the [fiftystater package](https://cran.r-project.org/package=fiftystater) by William Murphy. In contrast, `urbnmapr`:

  - Uses shapefiles from the US Census Bureau
  - Converts the shapefile data to a `tibble` data frame
  - Adds various identifiers for merging
  - Includes a county-level shapefile 

## Getting data

There are two methods for accessing map data in `urbnmapr`. First, `urbnmapr` contains two `tibble` data frames, `states` and `counties`, for easily accessing simple state and county map data. 

The `states` and `counties` tibbles can be used with `geom_polygon()` and `coord_map()` to create base maps of the continental United States, with Alaska and Hawaii displayed as insets:

```{r blank-states}
library(tidyverse)
library(urbnmapr)

states %>%
  ggplot(aes(long, lat, group = group)) +
    geom_polygon(fill = "grey", color = "#ffffff", size = 0.25) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

```{r blank-counties}
counties %>%
  ggplot(aes(long, lat, group = group)) +
    geom_polygon(fill = "grey", color = "#ffffff", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

Second, the function `get_urbn_map()` accesses additional maps including "ccdf", which includes elided territories roughly positioned in the direction of their true location, and "territories_counties", which includes territories below the continuous 48 states. 

```{r blank-territories-counties}
ccdf <- get_urbn_map(map = "territories_counties")

ccdf %>%
  ggplot(aes(long, lat, group = group)) +
    geom_polygon(fill = "grey", color = "#ffffff", size = 0.25) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

## Merging data

All tibbles include various identifiers to simplify merging data. The states `states` tibble contains `state_fips`, `state_abbv`, and `state_name`. The `counties` tibble contains `county_fips`, `state_abbv`, `state_fips`,  `county_name`, and `state_name`. Both tibbles can be piped into `ggplot2` to create a choropleth map.

```{r us-choropleth}
statedata %>% 
  left_join(states, by = "state_name") %>% 
  ggplot(mapping = aes(long, lat, group = group, fill = horate)) +
  geom_polygon(color = "#ffffff", size = 0.25) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Homeownership rate")
```

```{r county}
household_data <- left_join(countydata, counties, by = "county_fips") 

household_data %>%
  ggplot(aes(long, lat, group = group, fill = medhhincome)) +
    geom_polygon(color = "#ffffff", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

## Labels

Like the map data, there are two methods for accessing map labels in `urbnmapr`. First, `urbnmapr` contains two `tibble` data frames, `states_labels` and `counties_labels`, for easily labeling the corresponding maps for `states` and `counties`. 

```{r blank-states-labels}
library(tidyverse)
library(urbnmapr)

states %>%
  ggplot() +
    geom_polygon(aes(long, lat, group = group), 
                 fill = "grey", color = "#ffffff", size = 0.25) +
    geom_text(data = states_labels, aes(long, lat, label = state_abbv), size = 3) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

Second, the function `get_urbn_labels()` accesses additional labels including "ccdf", which includes elided territories roughly positioned in the direction of their true location, and "territories_counties", which includes territories below the continuous 48 states. 

```{r blank-ccdf-labels}
ccdf <- get_urbn_map(map = "ccdf")
ccdf_labels <- get_urbn_labels(map = "ccdf")

ccdf %>%
  ggplot() +
    geom_polygon(aes(long, lat, group = group), 
                 fill = "grey", color = "#ffffff", size = 0.25) +
    geom_text(data = ccdf_labels, aes(long, lat, label = state_abbv), size = 3) +  
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

## Styles

`library(urbnmapr)` works well with the [library(urbnthemes)](https://github.com/UI-Research/urbnthemes)

```{r theme-state}
library(urbnthemes)

set_urbn_defaults(style = "map")

statedata %>% 
  left_join(states, by = "state_name") %>% 
  ggplot(mapping = aes(long, lat, group = group, fill = horate)) +
  geom_polygon(color = "#ffffff", size = 0.25) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradientn(labels = scales::percent) +
   theme(legend.position = "right",
         legend.direction = "vertical",
         legend.title = element_text(face = "bold", size = 11),
         legend.key.height = unit(.2, "in")) +
  labs(fill = "Homeownership rate")

```

```{r theme-counties}
household_data %>%
  filter(state_name %in% c("Virginia", "Maryland", "District of Columbia")) %>%
  ggplot(aes(long, lat, group = group, fill = medhhincome)) +
    geom_polygon(color = "#ffffff", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    scale_fill_gradientn(labels = scales::dollar) +
    theme(legend.position = "right",
          legend.direction = "vertical",
          legend.title = element_text(face = "bold", size = 11),
          legend.key.height = unit(.25, "in")) +
  labs(fill = "Median household income")
```
