---
title: "Introducing urbnmapr"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `urbnmapr` package provides state and county shapefiles in `tibble` format 
that is compatible to map with `ggplot2`. 

Shapefiles include Alaska and Hawaii, transformed to be displayed as insets 
within the continental United States.

This work is heavily inspired by and derived in part from the fiftystater 
package <https://cran.r-project.org/package=fiftystater> by William Murphy.

## Usage

`urbnmapr` contains two `tibble` dataframes:

  - `states`
  - `counties`

The `states` and `counties` tibbles can be used with `geom_polygon()` and 
`coord_map()` to create base maps of the continental United States, with Alaska 
and Hawaii displayed as insets:

```{r blank-states}
library(ggplot2)
library(dplyr)
library(tibble)
library(magrittr)
library(urbnmapr)

states %>%
  ggplot(aes(long, lat, group = group)) +
   geom_polygon(fill = "grey", color = "#ffffff", size = 0.25) +
   coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

```{r blank-counties}
counties %>%
  ggplot(aes(long, lat, group = group)) +
   geom_polygon(fill = "grey", color = "#ffffff", size = 0.25) +
   coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

## Merging Data

The `states` and `counties` tibbles include various identifiers to make merging 
data easy.

```{r us-choropleth}
USArrests %>%
  rownames_to_column('state_name') %>%
  select(state_name, Murder) %>%
  left_join(states, by = 'state_name') %>%
  ggplot(aes(long, lat, group = group, fill = Murder)) +
   geom_polygon(color = "#ffffff") +
   coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

